# Reproducible Research Assignement 1 
___
```{r}
library(ggplot2)
library(Hmisc)
```
###Loading and preprocessing the data  
**1.Load the data (i.e. read.csv())**  
```{r}
unzip('activity.zip')
dat<-read.csv("activity.csv")
```
  
###What is mean total number of steps taken per day?  
**1.Calculate the total number of steps taken per day** 
```{r}
no_of_steps<-tapply(dat$steps, dat$date, sum, na.rm=TRUE)
```
  
**2.Make a histogram of the total number of steps taken each day** 
```{r}
qplot(no_of_steps, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)

```  
    
**3.Calculate and report the mean and median total number of steps taken per day**
````{r}
stepsByDayMean <- mean(no_of_steps)
stepsByDayMedian <- median(no_of_steps)
cat("Mean:",stepsByDayMean)
cat("Median:",stepsByDayMean)
```
  
  
###What is the average daily activity pattern?  
  
**1.Make a time series plot**
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=dat$steps),by=list(interval=dat$interval), FUN=mean, na.rm=TRUE)
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +geom_line() +xlab("5-minute interval") +ylab("average number of steps taken") 

```  
  
**2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**
```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2",averageStepsPerTimeBlock[mostSteps,'interval'])
cat("The Interval is:",timeMostSteps)
```
  
###Imputing missing values
  
**1.Calculate and report the total number of missing values in the dataset**
```{r}
missVal <- length(which(is.na(dat$steps)))
cat("Number of missing values:",missVal)
```
  
**2.Devise a strategy for filling in all of the missing values in the dataset.**
   
**3.Create a new dataset that is equal to the original dataset but with the missing data filled in.**
```{r}

activityDataImputed <- dat
activityDataImputed$steps <- impute(dat$steps, fun=mean)
```
  
**4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.**
```{r}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
cat("Mean:",stepsByDayMeanImputed)
cat("Median:",stepsByDayMedianImputed)
```
  
###Are there differences in activity patterns between weekdays and weekends?

**1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.**  
```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6),'weekend', 'weekday')

```
  
**2.Make a panel plot containing a time series plot**
```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + geom_line() + facet_grid(dateType ~ .) +xlab("5-minute interval") + ylab("avarage number of steps")

```
